---
title: "HK Immigration Passenger Traffic"
output: 
    rmarkdown::github_document:
        html_preview: false
---

## Data Source

On each day, the [Hong Kong Immigration Department](https://immd.gov.hk/eng/index.html) will update the passenger traffic at different control point of the previous day. For example, the passenger traffic statistics on `2021-12-16` can be found from the url below.

```
https://www.immd.gov.hk/eng/stat_20211216.html
```

By changing the date in the above url, the passenger traffic data of a particular date can be found.

## Data Extraction With Python

The function below takes a date value and return the "Passenger Traffic Statistic" on that day from the Immigration Department website. 

```{python python_function}
import requests
from bs4 import BeautifulSoup as bs
import pandas as pd
from datetime import datetime
from datetime import date
from datetime import timedelta
import re
import matplotlib.pyplot as plt

def getImmigrationStatDaily(d):
    url = "https://www.immd.gov.hk/eng/stat_{}.html".format(d.strftime("%Y%m%d"))
    response = requests.get(url)
    soup = bs(response.content, "html.parser")
    tr = soup.select("tbody tr")
    pattern = re.compile(r"(Arrival|Departure)")
    rows = list()
    for i in range(len(tr)):
        td = tr[i].select("td")
        hasHeader = list(map(lambda y: "headers" in set(y.attrs.keys()), td))
        td = [td[j] for j, x in enumerate(hasHeader) if x]
        headers = [cell.get_attribute_list("headers")[0] for cell in td]
        value = [cell.text for cell in td]
        value = [int(x.replace(",", "")) if bool(pattern.search(headers[i])) else x for i, x in enumerate(value)]
        rows.append(dict(zip(headers, value)))
    result = pd.DataFrame(rows)
    result[["Date"]] = d
    return(result)

df = getImmigrationStatDaily(date(2021,12,16))
```

```{r python_function_example, echo = FALSE}
knitr::kable(reticulate::py$df)
```


## Data Extraction With R 

The function below takes a date value and return the "Passenger Traffic Statistic" on that day from the Immigration Department website. 

```{r r_function}
getImmigrationStatDaily <- function(d) {
  url <- sprintf("https://www.immd.gov.hk/eng/stat_%s.html", format(d, "%Y%m%d"))
  response <- httr::content(httr::GET(url), encoding = "UTF-8")
  tr <- rvest::html_nodes(response, "tbody tr")
  result <- as.data.frame(t(sapply(tr, function(x) {
    y <- rvest::html_nodes(x, "td")
    y <- y[grep("headers", y)]
    r <- rvest::html_text(y)
    names(r) <- rvest::html_attr(y, "headers")
    return(r)
  })))
  cols <- grep("(Arrival|Departure)$", names(result))
  result[, cols] <- as.data.frame(lapply(result[, cols], function(x) as.numeric(gsub("[^0-9]", "", x))))
  if (ncol(result) > 1)	result$Date <- d else stop(paste0("No Data on ", d))
  return(result)
}

df <- getImmigrationStatDaily(as.Date("2021-12-16"))
```

```{r r_function_example}
knitr::kable(df)
```

## Airport Passenger Traffic Using Python

Below shows an example of extracting the airport passenger traffic using Python. 

```{python python_airport_extraction}
startdate = date(2020,7,1)
enddate = date.today() - timedelta(days = 1)
dates = [startdate + timedelta(days = i) for i in range((enddate - startdate).days + 1)]
dfs = [getImmigrationStatDaily(x) for x in dates]
df = pd.concat(dfs).reset_index(drop = True)
```

```{python python_airport_plotdf}
airportTraffic = df[(df["Date"] >= dates[0]) & (df["Control_Point"] == "Airport")][["Hong_Kong_Residents_Arrival", "Hong_Kong_Residents_Departure"]]
airportTraffic = airportTraffic.reset_index(drop = True)
airportTraffic["NetChange"] = airportTraffic["Hong_Kong_Residents_Arrival"] - airportTraffic["Hong_Kong_Residents_Departure"]

yearMonths = [str(x.year) + "-" + str(x.month).zfill(2) for x in dates]
airportTrafficMonthly = airportTraffic.groupby(yearMonths).sum()
airportTrafficMonthly["Month"] = list(airportTrafficMonthly.index)
airportTrafficMonthly = airportTrafficMonthly.reset_index(drop=True)
airportTrafficMonthly = airportTrafficMonthly[["Month", "Hong_Kong_Residents_Arrival", "Hong_Kong_Residents_Departure", "NetChange"]]
Total = airportTrafficMonthly.sum(numeric_only=True).append(pd.Series("Total"))
Total.index = list(Total.index)[0:3] + ["Month"]
outdf = airportTrafficMonthly.append(Total, ignore_index=True)
```

```{r python_airport_plot, echo = FALSE}
knitr::kable(reticulate::py$outdf)
```

```{python python_airport_linechart}
plt.plot(list(airportTrafficMonthly["Month"]), airportTrafficMonthly["Hong_Kong_Residents_Arrival"], label = "Arrival");
plt.plot(list(airportTrafficMonthly["Month"]), [-x for x in airportTrafficMonthly["Hong_Kong_Residents_Departure"]], label = "Departure");
plt.plot(list(airportTrafficMonthly["Month"]), airportTrafficMonthly["NetChange"], label = "Net Change");
plt.xticks(list(airportTrafficMonthly["Month"]), list(airportTrafficMonthly["Month"]), rotation='vertical');
plt.legend(loc = "lower center", bbox_to_anchor=(0.5, -0.4), ncol = 3);
plt.title("HK Passenger Traffer (Airport)");
plt.xlabel("Month");
plt.ylabel("Count");
plt.grid(axis = "y", linestyle='--', color = "#7c7c7c");
plt.axhline(0, color='k');
plt.tight_layout();
plt.show()
```

## Airport Passenger Traffic Using R

Below shows an example of extracting the airport passenger traffic using R. 

```{r r_airport_extraction}
dates <- seq(as.Date("2020-07-01"), Sys.Date() - 1, by = 1)
df <- plyr::rbind.fill(lapply(dates, getImmigrationStatDaily))
```

```{r r_airport_plotdf}
airportTraffic <- dplyr::filter(df, df$Control_Point == "Airport")[, c("Hong_Kong_Residents_Arrival", "Hong_Kong_Residents_Departure")]
airportTraffic$NetChange <- airportTraffic$Hong_Kong_Residents_Arrival - airportTraffic$Hong_Kong_Residents_Departure

yearMonths <- format(dates, "%Y-%m")
airportTrafficMonthly <- aggregate(airportTraffic, list(Month = yearMonths), sum)
knitr::kable(rbind(airportTrafficMonthly, as.data.frame(c(list(Month = "Total"), colSums(airportTrafficMonthly[, -1])))))
```

```{r r_airport_plot}
library(ggplot2)

plotdf <- airportTrafficMonthly
names(plotdf) <- c("Month", "Arrival", "Departure", "NetChange")
plotdf$Departure <- -plotdf$Departure
plotdf <- tidyr::gather(plotdf, Type, Count, Arrival:NetChange)

ggplot(data = plotdf, aes(x = Month, y = Count, group = Type)) + 
    geom_line(aes(color = Type)) + 
    geom_point(aes(color = Type)) + 
    ggtitle("HK Passenger Traffer (Airport)") + 
    theme(
        legend.position="bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust=0.5)
    )
```

